Relationship between solar wind, $D_{st}$, and plasmasphere mass density on one-hour time scales

### 2015-06-25 paper.tex
Updated figures. Fixing text

\cite{Takahashi2010} state that spikes in the Disturbance Storm Time ($D_{st}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{st}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{st}$ measurements.

@@

In this work events are considered when $D_{st}$ or $\rho_{eq}$ crosses a threshold value.  In finding events, all fill values were replaced with linearly interpolated values.

@@

\subsection{D$_{st}$ Events}
Two event indicators are looked at in this study. The first is looking for a drop in DST below the threshhold of $-50nT$ specified in \cite{Takahashi2010}, dubbed the "onset", and then considering the timeframe a event until $D_{st}$ passes back above the $-50nT$ threshold. This method finds 669 such periods between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours. A subset of longer duration events will be looked at later. Figure \ref{AllData} shows the average values of $B_Z$, $V_{SW}$, $D_{st}$, $F_{10.7}$, and Mass Density for a period of 24 hours before and 48 hours after a event onsppet, as well as a plot of mass density with the found event times highlighted.

@@

\subsection{Mass Density Events}
Figure \ref{MassStorm} shows this same algorithm, but looking for a rise in mass density over a value of 40$g/cm^3$. This results in 130 events with a mean duration of 32 hours and a median duration of 17 hours, marked in red on the left figure.

@@

Similarly, looking at the 1 day averaged data with and without the $F_{10.7}$ dependence, Figure \ref{nof107day} shows that the results of \cite{Takahashi2010} hold only once the dependence has been removed. The 27 day trend couldn't be determined as not enough valid data existed to test it.

@@

@@

% Change x-label to be ``Time from start of event (hr)'' and put vertical lines in each panel at t = 0 (somewhere I have a code called vertline.m)
\begin{figure}[htp]
\centering
(a)
\includegraphics[scale=0.7]{paperfigures/massdensitystorms.eps}
(b)
\includegraphics[scale=0.7]{paperfigures/dststorms.eps}
\caption{Averages around time of (a) 669 $\rho_{eq}$ and (b) 130 $D_{st}$ events along with thresholds used to define an event. The number of values used in the averages for mass density is shown in the bottom panel (a) and (b).}
\label{DSTstorm}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dst.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-mass.eps}
\caption{(a) Average of solar wind and near-Earth measurements around the time $D_{st}$ crossed below -40~nT onset. (b) Same as (a) except around time intervals where mass density crossed above $40 amu/cm^3$.  The bottom panel shows the number of mass density data points that were used for the average.}
\label{MassStorm}
\end{figure}
\clearpage

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.7]{paperfigures/stormavs-diffden-10amu.png}
\includegraphics[scale=0.7]{paperfigures/stormavs-diffden-30percent.png}
\caption{diff($\rho_{eq}$) $>$ 20$\frac{amu}{hour}$, diff($\rho_{eq}$) $>$ 30$\frac{\%}{hour}$}
\label{rhochange}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.7]{paperfigures/stormavs-dst-nof107.png}
\includegraphics[scale=0.7]{paperfigures/f107removed.png}
\caption{$\rho_{eq}$ with $F_{10.7}$ removed, and Difference from original}
\label{nof107}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.7]{paperfigures/stormavs-dst-day.png}
\includegraphics[scale=0.7]{paperfigures/stormavs-dst-day-nof107.png}
\caption{Original $\rho_{eq}$, and with $F_{10.7}$ dependence removed, both at 1-day averages}
\label{nof107day}
\end{figure}

>>>>

\cite{Takahashi2010} state that spikes in the Disturbance Storm Time ($D_{st}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{st}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{st}$ measurements. Figure \ref{DailyAverages} show that the trends observed by \cite{Takahashi2010} hold true in this analysis as well.

@@

In this work events are considered when $D_{st}$ or $\rho_{eq}$ crosses a threshold value, as indicated in \ref{StormCutoff}.  In finding events, all fill values were replaced with linearly interpolated values.  Figure \ref{AllData} shows the values of $B_Z$, $V_{SW}$, $D_{st}$, $F_{10.7}$, and Mass Density for the duration of the dataset.

@@

\subsection{D$_{st}$ Events}
Two event indicators are looked at in this study. The first is looking for a drop in DST below the threshhold of $-40nT$ specified in \cite{Takahashi2010}, dubbed the "onset", and then considering the timeframe an event until $D_{st}$ passes back above the $-40nT$ threshold. This method finds 669 such periods between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours. Figure \ref{Storms} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{Storms}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40nT$. The dashed red lines indicate plus and minus one standard deviation of values from all events that went into the average. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other variables in the stack. A subset of longer duration events will be looked at later.

@@

\subsection{Mass Density Events}
Figure \ref{Storms}b shows this same algorithm, but looking for a rise in mass density over a value of 40$g/cm^3$. This results in 130 events with a mean duration of 32 hours and a median duration of 17 hours, marked in red on the left figure.

@@

\section{Appendix}
\subsection{Bias}

While the GOES Satellites tend to keep an average geostationary distance of around $L=6.8R_E$ shown in \cite{Takahashi2010}, the actual plasmapause location varies significantly , as shown in \cite{OBrien2003}. This means that during periods of large $D_{st}$ the plasmapause may be far from the point of measurement of $\rho_{eq}$ creating a discord in the correlation. \cite{Gallagher2000} provides a model for $\rho_{eq}$ at a range of L-shells and a brief discussion of the elemental contributions.

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dst-day.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-dst-27day.eps}
\caption{One-day and 27-day averaged $D_{st}$ storms}
\label{DailyAverages}
\end{figure}
\clearpage

@@

% Change x-label to be ``Time from start of event (hr)'' and put vertical lines in each panel at t = 0 (somewhere I have a code called vertline.m)
\begin{figure}[htp]
\centering
(a)
\includegraphics[scale=0.7]{paperfigures/massdensitystorms.eps}
(b)
\includegraphics[scale=0.7]{paperfigures/dststorms.eps}
\caption{Averages around time of (a) 669 $\rho_{eq}$ and (b) 130 $D_{st}$ events along with thresholds used to define an event. The number of values used in the averages for mass density is shown in the bottom panel (a) and (b).}
\label{StormCutoff}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dst.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-mass.eps}
\caption{(a) Average of solar wind and near-Earth measurements around the time $D_{st}$ crossed below -40~nT onset. (b) Same as (a) except around time intervals where mass density crossed above $40 amu/cm^3$.  The bottom panel shows the number of mass density data points that were used for the average.}
\label{Storms}
\end{figure}
\clearpage

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-diffden-10amu.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-diffden-30percent.eps}
\caption{diff($\rho_{eq}$) $>$ 10$\frac{amu}{hour}$, diff($\rho_{eq}$) $>$ 30$\frac{\%}{hour}$}
\label{rhochange}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dst-nof107.eps}
\includegraphics[scale=0.6]{paperfigures/f107removed.eps}
\caption{$D_{st}$ events where $\rho_{eq}$ has $F_{10.7}$ dependence removed, and Difference between original $\rho_{eq}$ and $F_{10.7}$-less $\rho_{eq}$}
\label{nof107}
\end{figure}
\clearpage
